Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period

### 2019-12-10 analysis/paper/paper.Rmd
added one reference and editted the conclusion section

The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century. A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh\\]. The origin of light gray glazed jars was thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007]. Regarding large dark brown glazed stoneware jars, they might be made in Southeast Asia and are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages. The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131]. Those jars might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan. Either way indicates that the Europeans played an important role in introducing those jars to Taiwan. We used the presence and absence of light grey glazed jar and large stoneware jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and the start of the European influence at Kiwulan.

@@

Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by colonial activities or involved in the colonial economy centred in the major European colonies [@Trabert2017]. Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks. Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation. The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods. The greater amount and diversity of ornament types during the European period reflects international exchanges in a colonial context. Ornaments were viewed as prestige goods in the local indigenous culture before European contact. After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signalling wealth and privileged trading connections among the inhabitants of Kiwulan. These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality.

We are still far from understanding the full variety of colonial impacts on peripheral indigenous communities. The Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations. Ornaments give insights into the amplification of social inequality stimulated by European colonisation. It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences. Future work could extend this approach to studies of other trade goods such as ceramics. We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardisation, and mortuary practices.

>>>>

The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century. A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh1995]. The origin of light gray glazed jars was thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007]. Regarding large dark brown glazed stoneware jars, they might be made in Southeast Asia and are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages. The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131]. Those jars might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan. Either way indicates that the Europeans played an important role in introducing those jars to Taiwan. We used the presence and absence of light grey glazed jar and large stoneware jars, together with the radiocarbon dates, to identify excavation units associated with the pre-European and the start of the European influence at Kiwulan.

@@

Examination of the indirect influence of colonialism can reveal how peripheral areas were influenced by colonial activities or involved in the colonial economy centred in the major European colonies [@Trabert2017]. Kiwulan in northeastern Taiwan is an exceptional case study as an East Asian location that was relatively isolated and peripheral, and yet connected by regional and global trade networks. Kiwulan provides valuable insights into the discussion of indirect colonial influence on local societies living beyond the borders of direct European colonial occupation. The frequency and spatial distribution of body ornaments at Kiwulan present three distinct patterns during different dominant culture interaction periods. The greater amount and diversity of ornament types during the European period reflects an increasing use in ornaments in a colonial context. Before European contact, ornaments were traded into local indigenous societies via the regional exchagne network with Chinese merchants and viewed as prestige goods in local indigenous culture according to their distribution in the archaeological contexts. After the arrival of the Europeans, the exotic and powerful image carried by those ornaments may have intensified, further signalling wealth and privileged trading connections among the inhabitants of Kiwulan. These symbolic values may have stimulated more competition between aggrandizing individuals for prestige and wealth accumulation at Kiwulan, which might have resulted in an increase in social inequality. This might also indicate an act of intentional resistance to the intrusion of the Europeans by using more ornaments that is part of culture tradition.

We are still far from understanding the full variety of colonial impacts on peripheral indigenous communities. The origin of ornaments and how their sources changed over time would give more information to construct a clear picture of complex trade networks during this periods. By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations. Ornaments give insights into the amplification of social inequality stimulated by European colonisation. It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences. Future work could extend this approach to studies of other trade goods such as ceramics. We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardisation, and mortuary practices.
